Data Exploration: IUCN and AquaMaps

### 2015-12-21 data_explore/data_explore.Rmd
updated to ggplot2 2.0 and redid plots

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  geom_bar(stat = 'bin', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 2, color = 'grey20') +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Relative proportion of species per quadrant', 
       title = 'Taxonomic group by quadrant',
       fill = 'Alignment')

@@

### Plot the bar chart
barchart_spp_risk_quads <- ggplot(spp_risk_quadrants %>% 
                                  transform(popn_category = reorder(popn_category, cat_rank)), 
                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  geom_bar(stat = 'bin', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 2, color = 'grey20') +
  coord_flip() +
  labs(x = 'Risk category', 
       y = 'Relative proportion of species per quadrant', 
       title = 'Risk category by quadrant',
       fill = 'Alignment')

>>>>

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  geom_bar(stat = 'count', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 2, color = 'grey20') +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Relative proportion of species per quadrant', 
       title = 'Taxonomic group by quadrant',
       fill = 'Alignment')

@@

### Plot the bar chart
barchart_spp_risk_quads <- ggplot(spp_risk_quadrants %>% 
                                  transform(popn_category = reorder(popn_category, cat_rank)), 
                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  geom_bar(stat = 'count', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 2, color = 'grey20') +
  coord_flip() +
  labs(x = 'Risk category', 
       y = 'Relative proportion of species per quadrant', 
       title = 'Risk category by quadrant',
       fill = 'Alignment')

### 2015-12-21 data_explore/data_explore_am_thresh.Rmd
updated to ggplot2 2.0 and redid plots

```{r aquamaps probabilities, warning = FALSE, message = FALSE, cache.lazy = FALSE, cache = TRUE, verbose = FALSE, results = FALSE, echo = FALSE, include = FALSE}

@@

# spp_cells <- fread(input = file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), verbose = FALSE, header = TRUE, 
#                   colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec

@@

hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  geom_vline(x = spp_df3_mean, color = 'darkred', linetype = 'dashed') + 
  annotate('text', x = spp_df3_mean, y = .16, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'mean') + 
  geom_vline(x = spp_df3_med, color = 'grey30', linetype = 'dashed') + 
  annotate('text', x = spp_df3_med, y = .16, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'median') + 
  scale_fill_manual(values = c('LC' = 'green4', 
                               'NT' = 'greenyellow', 
                               'VU' = 'orange', 
                               'EN' = 'orangered2', 
                               'CR' = 'red4',
                               'DD' = 'grey60',
                               'NE' = 'grey70')) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)))

@@

hist_area_vs_threshold2 <- ggplot(spp_df3a, aes(prop, fill = iucn_code)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  geom_vline(x = spp_df3a_mean, color = 'darkred', linetype = 'dashed') + 
  annotate('text', x = spp_df3a_mean, y = .16, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'mean') + 
  geom_vline(x = spp_df3a_med, color = 'grey30', linetype = 'dashed') + 
  annotate('text', x = spp_df3a_med, y = .16, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'median') + 
  scale_fill_manual(values = c('LC' = 'green4', 
                               'NT' = 'greenyellow', 
                               'VU' = 'orange', 
                               'EN' = 'orangered2', 
                               'CR' = 'red4',
                               'DD' = 'grey60',
                               'NE' = 'grey70')) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3a)))

@@

scatter_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = mean_prop)) + 
  ggtheme_plot + 
  geom_point(aes(y = median_prop), color = 'grey60') +
  geom_point() + 
  geom_abline(slope = -1, intercept = 1, color = 'red') + 
  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = 'Mean (median) species range after presence threshold cutoff',
       x = 'Presence threshold cutoff', 
       y = 'Mean (median) relative species range')

@@

  geom_ribbon(aes(ymax = quant91, ymin = quant09), alpha = .5, color = NA, fill = 'grey80') +
  geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = 'grey70') +
  geom_point(color = 'grey30') + 
  geom_point(aes(y = mean_prop), color = 'darkred') + 
  geom_abline(intercept = area_mean_model[['coefficients']][1],
              slope     = area_mean_model[['coefficients']][2],
              color = 'red') + 
  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
  geom_vline(x = 0.50, color = 'blue', linetype = 'dashed') +
  geom_vline(x = 0.60, color = 'blue', linetype = 'dashed') +
  annotate('text', x = .40, y = .05, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'Ocean Health Index: 40%') + 
  annotate('text', x = .50, y = .05, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'MPA Gap Analysis: 50%') + 
  annotate('text', x = .60, y = .05, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'Marine mammals: 60%') + 
  annotate('text', x = .75, y = .85,
           size = 2, color = 'grey20',
           label = 'R^2 = .995') + 
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = 'Relative species range after presence threshold',
       x = 'Presence threshold cutoff', 
       y = 'Mean (median) relative species range')

>>>>

```{r aquamaps_load_data, warning = FALSE, message = FALSE, echo = FALSE, include = FALSE}

@@

@@

hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  geom_vline(xintercept = spp_df3_mean, color = 'darkred', linetype = 'dashed') + 
  annotate('text', x = spp_df3_mean, y = .16, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'mean') + 
  geom_vline(xintercept = spp_df3_med, color = 'grey30', linetype = 'dashed') + 
  annotate('text', x = spp_df3_med, y = .16, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'median') + 
  scale_fill_manual(values = c('LC' = 'green4', 
                               'NT' = 'greenyellow', 
                               'VU' = 'orange', 
                               'EN' = 'orangered2', 
                               'CR' = 'red4',
                               'DD' = 'grey60',
                               'NE' = 'grey70')) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)))

@@

hist_area_vs_threshold2 <- ggplot(spp_df3a, aes(prop, fill = iucn_code)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  geom_vline(xintercept = spp_df3a_mean, color = 'darkred', linetype = 'dashed') + 
  annotate('text', x = spp_df3a_mean, y = .16, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'mean') + 
  geom_vline(xintercept = spp_df3a_med, color = 'grey30', linetype = 'dashed') + 
  annotate('text', x = spp_df3a_med, y = .16, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'median') + 
  scale_fill_manual(values = c('LC' = 'green4', 
                               'NT' = 'greenyellow', 
                               'VU' = 'orange', 
                               'EN' = 'orangered2', 
                               'CR' = 'red4',
                               'DD' = 'grey60',
                               'NE' = 'grey70')) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3a)))

@@

scatter_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = mean_prop)) + 
  ggtheme_plot + 
  geom_point(aes(y = median_prop), color = 'grey60') +
  geom_point() + 
  geom_abline(slope = -1, intercept = 1, color = 'red') + 
  geom_vline(xintercept = 0.40, color = 'blue', linetype = 'dashed') +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = 'Mean (median) species range after presence threshold cutoff',
       x = 'Presence threshold cutoff', 
       y = 'Mean (median) relative species range')

@@

  geom_ribbon(aes(ymax = quant91, ymin = quant09), alpha = .5, color = NA, fill = 'grey80') +
  geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = 'grey70') +
  geom_point(color = 'grey30') + 
  geom_point(aes(y = mean_prop), color = 'darkred') + 
  geom_abline(intercept = area_mean_model[['coefficients']][1],
              slope     = area_mean_model[['coefficients']][2],
              color = 'red') + 
  geom_vline(xintercept = 0.40, color = 'blue', linetype = 'dashed') +
  geom_vline(xintercept = 0.50, color = 'blue', linetype = 'dashed') +
  geom_vline(xintercept = 0.60, color = 'blue', linetype = 'dashed') +
  annotate('text', x = .40, y = .05, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'Ocean Health Index: 40%') + 
  annotate('text', x = .50, y = .05, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'MPA Gap Analysis: 50%') + 
  annotate('text', x = .60, y = .05, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'Marine mammals: 60%') + 
  annotate('text', x = .75, y = .85,
           size = 2, color = 'grey20',
           label = 'R^2 = .995') + 
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = 'Relative species range after presence threshold',
       x = 'Presence threshold cutoff', 
       y = 'Mean (median) relative species range')
